Acceleration of Global Search by Implementing Dual Estimates for Lipschitz Constant\thanks{This research was supported by the the Russian Science Foundation, project No.\,16-11-10150.}

### 2019-05-22 2019/NUMTA_Barkalov/LNCS/numta19.tex

\begin{document}
%
\title{Acceleration of Global Search through Dual Lipschitz Constant Estimates\thanks{This research was supported by the the Russian Science Foundation, project No.\,16-11-10150.}}
%
\titlerunning{Acceleration of Global Search}
% If the paper title is too long for the running head, you can set
% an abbreviated paper title here
%
\author{Roman Strongin%\orcidID{0000-0003-0390-6695} 
\and Konstantin Barkalov%\orcidID{0000-0001-5273-2471} 
\and Semen Bevzuk%\orcidID{0000-0002-3845-5356}
}
%
\authorrunning{R. Strongin et al.}
% First names are abbreviated in the running head.
% If there are more than two authors, 'et al.' is used.
%
\institute{
Lobachevsky State University of Nizhni Novgorod, Nizhni Novgorod, Russia
\email{konstantin.barkalov@itmm.unn.ru}
}
%
\maketitle              % typeset the header of the contribution
%
\begin{abstract}
The paper considers global optimization problems with a black-box objective 
function satisfying the Lipschitz condition. Efficient algorithms for this 
class of problems require reliable estimates of the Lipschitz constant to be 
introduced. The various approaches have been proposed to take into account both
global and local properties of the objective function. In particular, algorithms
using local estimates of the Lipschitz constant have shown their potential.
The new approach presented in this paper is based on simultaneous use of two
estimates: one is substantially larger than the other. 
The larger estimate ensures global convergence and the smaller one reduces 
the total number of trials needed to find the global optimizer.
Results of numerical experiments on the random sample of multidimensional 
functions demonstrate the efficiency of the suggested approach.

@@

The theory of convergence of this algorithm is provided in \cite{Strongin2000}.

@@

%Russian
The global search algorithm presented in previous section is intended for solving the multiextremal problems, in which the objective function satisfies the Lipschitz condition. It is not necessary to define the value of the constant for the algorithm convergence. The estimation of the constant is performed in the course of global search on the base of available search information. 
According to the theorem in \cite{Strongin2000}, the sequence of the trial points $\{y^k\}$ will converge to the global minimizer $y^*$ if the condition 
\begin{equation}\label{cond}
r\mu > 2^{3-1/N}L\sqrt{N+3}.
\end{equation}
\noindent is satisfied. Thus,  an appropriate choice of the parameter $r$ from (\ref{R}) allows using the value $r\mu / 2^{3-1/N}\sqrt{N+3}$ as an estimate of the Lipschitz constant for the objective function $\varphi(y)$.

@@

New rule of calculating the characteristic $R(i)$ of the interval $(x_{i-1}, x_i)$ will consist of the following operations:
\begin{itemize}
\item
Calculate the value $R_{glob}(i)$ corresponding to the larger estimate of the Lipschitz constant
\[
R_{glob}(i)=\Delta_i+\frac{(z_i-z_{i-1})^2}{r_{glob}^2\mu^2\Delta_i}-2\frac{z_i+z_{i-1}-2z^*}{r_{glob}\mu}.
\]
\item
Calculate the value $R_{loc}(i)$ corresponding to the smaller estimate of the Lipschitz constant
\[
R_{loc}(i)=\Delta_i+\frac{(z_i-z_{i-1})^2}{r_{loc}^2\mu^2\Delta_i}-2\frac{z_i+z_{i-1}-2z^*}{r_{loc}\mu}.
\]
\item
Determine the characteristic $R(i)$ as
\begin{equation}\label{pho}
R(i) = \max\{\rho R_{loc}(i),R_{glob}(i)\}, where \rho = \left(\frac{1-1/r_{glob}}{1-1/r_{loc}}\right)^2,
\end{equation}
and fix the value $r = r_{loc}$ if $\rho R_{loc}(i) > R_{glob}(i)$, otherwise fix $r=r_{glob}$.
\item
This value of $r$ is used in Rule 5 of the algorithm in the computing of the next trial point.   
\end{itemize}.

@@

%Russian
Each series of problems has been solved by the original global search algorithm (GSA) and by the method with two estimates of the Lipschitz constant (GSA-DL). The evolvent constructed using the parameter $m = 10$ was used for the dimensionality reduction in both algorithms. The relative accuracy of the solution search was $\delta = 0.01$ for $N=3,4$ and $\delta = 0.03$ for $N=5$.
The maximum allowable number of iterations per problem was $K_{max} = 10^6$.

The averaged numbers of iterations performed by the algorithms are presented in Table \ref{tab:1}.
At that, for GSA the values of parameter $r=4.8$ when solving the problems of Simple class and $r=5.6$ when solving the problems of Hard class were used. 
These values are the minimum ones (with the accuracy 0.1), at which all problems have been solved successfully.
When solving the problems from the above classes by GSA-DL, the value of parameter $r$ specified above was selected for the upper estimate of the Lipschitz constant, i.e. the value $r_{glob} = r$ was set, which was complemented by the values $r_{loc}=1.8$, $r_{loc}=2.1$ and $r_{loc}=2.4$.

\begin{table}
	\caption{Average number of iterations}
	\label{tab:1}
	\center
	\begin{tabular}{ccccccccc}
		\hline\noalign{\smallskip}
		 & \multicolumn{2}{c}{$N=3$} & & \multicolumn{2}{c}{$N=4$} & & \multicolumn{2}{c}{$N=5$}  \\
		\noalign{\smallskip} \cline{2-3} \cline{5-6} \cline{8-9} \noalign{\smallskip}
		 & \textit{Simple} & \textit{Hard} & & \textit{Simple} & \textit{Hard} & & \textit{Simple} & \textit{Hard}  \\
		\noalign{\smallskip} \hline \noalign{\smallskip}
									AGS	&	2444	&	5345	& &	28415	&	77470	& &		&		\\
AGS-DL, $r_{loc}=1.8$	&	1372	&	2632	& &	13273	&	37715	& &		&		\\
AGS-DL, $r_{loc}=2.1$	&	1502	&	2805	& &	14826	&	38843	& &		&		\\
AGS-DL, $r_{loc}=2.4$	&	1567	&	2868	& &	19447	&	40342	& &		&		\\
		\noalign{\smallskip}\hline
	\end{tabular}
\end{table}

>>>>

\begin{document}
%
\title{Acceleration of Global Search by Implementing Dual Estimates for Lipschitz Constant\thanks{This research was supported by the the Russian Science Foundation, project No.\,16-11-10150.}}
%
\titlerunning{Acceleration of Global Search}
% If the paper title is too long for the running head, you can set
% an abbreviated paper title here
%
\author{Roman Strongin%\orcidID{0000-0003-0390-6695} 
\and Konstantin Barkalov%\orcidID{0000-0001-5273-2471} 
\and Semen Bevzuk%\orcidID{0000-0002-3845-5356}
}
%
\authorrunning{R. Strongin et al.}
% First names are abbreviated in the running head.
% If there are more than two authors, 'et al.' is used.
%
\institute{
Lobachevsky State University of Nizhni Novgorod, Nizhni Novgorod, Russia
\email{konstantin.barkalov@itmm.unn.ru}
}
%
\maketitle              % typeset the header of the contribution
%
\begin{abstract}
The paper considers global optimization problems with a black-box objective 
function satisfying the Lipschitz condition. Efficient algorithms for this 
class of problems require reliable estimates of the Lipschitz constant to be 
introduced. The various approaches have been proposed to take into account both
global and local properties of the objective function. In particular, algorithms
using local estimates of the Lipschitz constant have shown their potential.
The new approach presented in this paper is based on simultaneous use of two
estimates: one is substantially larger than the other. 
The larger estimate ensures global convergence and the smaller one reduces 
the total number of trials needed to find the global optimizer.
Results of numerical experiments on the random sample of multidimensional 
functions demonstrate the efficiency of the suggested approach.

@@

The theory of convergence of this algorithm is provided in \cite{Strongin2000}. 
The algorithm is allows for an efficient
parallelization for shared and distributed memory \cite{Gergel2003}  and for accelerators \cite{Gergel2016}.

@@

%Russian
The global search algorithm presented in previous section is intended for solving the multiextremal problems, in which the objective function satisfies the Lipschitz condition. It is not necessary to define the value of the constant for the algorithm convergence. The estimation of the constant is performed in the course of global search on the base of available search information. 
According to the theorem from \cite{Strongin2000}, the sequence of the trial points $\{y^k\}$ will converge to the global minimizer $y^*$ if the condition 
\begin{equation}\label{cond}
r\mu > 2^{3-1/N}L\sqrt{N+3}.
\end{equation}
\noindent is satisfied. Thus,  an appropriate choice of the parameter $r$ from (\ref{R}) allows using the value $r\mu / 2^{3-1/N}\sqrt{N+3}$ as an estimate of the Lipschitz constant for the objective function $\varphi(y)$.

@@

New rule of calculating the characteristic $R(i)$ of the interval $(x_{i-1}, x_i)$ will consist of the following operations:
\begin{itemize}
\item
Calculate the value $R_{glob}(i)$ corresponding to the larger estimate of the Lipschitz constant
\[
R_{glob}(i)=\Delta_i+\frac{(z_i-z_{i-1})^2}{r_{glob}^2\mu^2\Delta_i}-2\frac{z_i+z_{i-1}-2z^*}{r_{glob}\mu}.
\]
\item
Calculate the value $R_{loc}(i)$ corresponding to the smaller estimate of the Lipschitz constant
\[
R_{loc}(i)=\Delta_i+\frac{(z_i-z_{i-1})^2}{r_{loc}^2\mu^2\Delta_i}-2\frac{z_i+z_{i-1}-2z^*}{r_{loc}\mu}.
\]
\item
Determine the characteristic $R(i)$ as
\begin{equation}\label{pho}
R(i) = \max\{\rho R_{loc}(i),R_{glob}(i)\}, \textrm{where} \; \rho = \left(\frac{1-1/r_{glob}}{1-1/r_{loc}}\right)^2,
\end{equation}
and fix the value $r = r_{loc}$ if $\rho R_{loc}(i) > R_{glob}(i)$, otherwise fix $r=r_{glob}$.
\item
This value of $r$ is used in Rule 5 of the algorithm in the computing of the next trial point.   
\end{itemize}.

@@

%Russian
Each series of problems has been solved by the original global search algorithm (GSA) and by the method with two estimates of the Lipschitz constant (GSA-DL). The evolvent constructed using the parameter $m = 10$ was used for the dimensionality reduction in both algorithms. The relative accuracy of the solution search was $\delta = 0.01$.% for $N=3,4$ and $\delta = 0.02$ for $N=5$. 
The maximum allowable number of iterations per problem was $K_{max} = 10^6$.

The averaged numbers of iterations performed by the algorithms are presented in Table \ref{tab:1}.
At that, for GSA the values of parameter $r=4.8$ when solving the problems of Simple class and $r=5.6$ when solving the problems of Hard class were used. 
These values are the minimum ones (with the accuracy 0.1), at which all problems have been solved successfully.
When solving the problems from the above classes by GSA-DL, the value of parameter $r$ specified above was selected for the upper estimate of the Lipschitz constant, i.e. the value $r_{glob} = r$ was set, which was complemented by the values $r_{loc}=1.8$, $r_{loc}=2.1$ and $r_{loc}=2.4$. 
The number of unsolved problems is specified in brackets.

\begin{table}
	\caption{Average number of iterations}
	\label{tab:1}
	\center
	\begin{tabular}{ccccccccc}
		\hline\noalign{\smallskip}
		 & \multicolumn{2}{c}{$N=3$} & & \multicolumn{2}{c}{$N=4$} & & \multicolumn{2}{c}{$N=5$}  \\
		\noalign{\smallskip} \cline{2-3} \cline{5-6} \cline{8-9} \noalign{\smallskip}
		 & \textit{Simple} & \textit{Hard} & & \textit{Simple} & \textit{Hard} & & \textit{Simple} & \textit{Hard}  \\
		\noalign{\smallskip} \hline \noalign{\smallskip}
									AGS	&	2444	&	5345	& &	28415	&	77470	& & 25220(1) & 126138(4)		\\
AGS-DL, $r_{loc}=1.8$	&	1372	&	2632	& &	13273	&	37715	& &	12702    & 94296(1)			\\
AGS-DL, $r_{loc}=2.1$	&	1502	&	2805	& &	14826	&	38843	& &	15213    & 90792(2)			\\
AGS-DL, $r_{loc}=2.4$	&	1567	&	2868	& &	19447	&	40342	& &	18239    & 106438(2)		\\
		\noalign{\smallskip}\hline
	\end{tabular}
\end{table}
